---
output:
  html_document: default
  pdf_document: default
---

#### Problem 1 (3 + 3 + 4 + 2)

#### (a) 

(3 points) Consider a binary classication problem with $Y\in\{-1,+1\}$ and two real-valued features, 
$X_1,X_2\in \mathbb{R}$. Suppose that we have learned a (Gaussian) naive Bayes classifier and obtained parameter estimates $\hat{\mu}_{-,j}= 0, \hat{\sigma}^{2}_{-,j}=1$ and $\hat{\mu}_{+,j}= 0, \hat{\sigma}^{2}_{-,j}=16$ for $j\in\{1,2\}$. Further, we use a uniform class prior $\hat{p}(y)=1/2$ for $y\in\{-1,+1\}$. 
Use the Bayes formula to compute the posterior probability
$$
P(Y = +1 | X_1 = 1,X_2 = 2)
$$

#### Answer:
Bayes formula:
$$
P(A|B) = \frac{P(A)P(B|A)}{P(B)}
$$
$$
P(Y = +1 | X_1 ,X_2) \\ \\
= \frac{P(Y = +1 )P(X_1,X_2|Y = +1 )}{P(X_1,X_2)} \\
= \frac{P(Y = +1 )P(X_1|Y = +1 )P(X_2|Y = +1 )}{P(X_1,X_2)} (Naive \  Bayes)\\
$$

Further calculation on paper lead to

$$
P(Y = +1 | X_1 ,X_2) = \frac{e^{\frac{-(x_1^2 + x_2^2)}{32}}}{e^{-\frac{-(x_1^2 + x_2^2)}{32}}+e^{\frac{-(x_1^2 + x_2^2)}{2}}}\\ \\
$$
Define this as R function probYplus1(x1,x2):
```{r}
probYplus1 <- function(x1,x2)
{
  px1 = 1/16*exp(-(x1^2+x2^2)/32)
  px2 = exp(-(x1^2+x2^2)/2)
  return(px1/(px1+px2))
}
probYminus1 <- function(x1,x2)
{
  px1 = 1/16*exp(-(x1^2+x2^2)/32)
  px2 = exp(-(x1^2+x2^2)/2)
  return(px2/(px1+px2))
}
```

```{r}
probYplus1(1,2)
probYminus1(1,2)
```
$$
P(Y = +1 | X_1=1 ,X_2=2) = \frac{e^{\frac{-(1 + 2)}{32}}}{e^{-\frac{-(1 + 2)}{32}}+e^{\frac{-(1 + 2)}{2}}} = 0.3944\\ 
$$
## (b) 
(3 points) Visualize the class posterior $P(Y = +1 | x)$, where $x = (x1, x2) contains the observed feature values on a suitable grid of points.
Hint: This is quite similar to Exercise 2.1.
```{r}
#create grid 
xgrid <- .25*(-20:20)
ygrid <- .25*(-20:20)
grid <- expand.grid(xgrid, ygrid)
py = probYplus1(grid[,1],grid[,2])
pyDraw <- matrix(py, nrow=41, ncol=41)

contour(xgrid, ygrid, pyDraw, xlab="x1", ylab="x2")
image(xgrid, ygrid, pyDraw, xlab="x1", ylab="x2")
persp(xgrid, ygrid, pyDraw, zlab="Probability y = +1", xlab="x1", ylab="x2",phi = 35, theta = 10)
```
